---
title: "Homework 3"
author: Allison Stewart
output: github_document
---

This is my solution for Homework 3. 

```{r}
library(tidyverse)
library(readxl)
library(dplyr)
library(ggplot2)
library(patchwork)
library(ggridges)
```

## Problem 1 

```{r}
library(p8105.datasets)
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns. Observations are at the level of items in orders by user. There are also item variables -- name, aisle, department, and some numeric codes. 

How many aisles, and which are most items from? 

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Make a plot! 

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(
    aisle = factor(aisle), 
    aisle = fct_reorder(aisle,n)
  ) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

Make a table! 

```{r}
instacart %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetable fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(aisle, rank) %>% 
  knitr::kable()
```

Apples vs. ice cream 

```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour
  )
```

## Problem 2 

```{r}
accel_df = 
  read.csv("./accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440, 
    names_to = "activity_minute", 
    values_to = "activity_count") %>% 
  relocate(day_id) %>% 
  mutate("weekday_weekend" = day) %>% 
  mutate(weekday_weekend = recode(weekday_weekend, Monday = "Weekday", Tuesday = "Weekday", Wednesday = "Weekday", Thursday = "Weekday", Friday = "Weekday", Saturday = "Weekend", Sunday = "Weekend")) %>% 
  mutate_at(vars(activity_minute), as.factor) %>% 
  mutate_at(vars(day), as.factor) %>% 
  mutate_at(vars(weekday_weekend), as.factor)
```

This dataset contains five weeks of accelerometer data collected on a 63-year-old male who was admitted to the Columbia University Medical Center and diagnosed with congestive heart failure. The variables in the dataset are `r variable.names(accel_df)`. The activity_minute variable represents each minute of a 24-hour day starting at midnight, with the respective activity counts for each minute shown by the activity_count variable. The final table contains `r nrow(accel_df)` rows and `r ncol(accel_df)` columns. The mean activity count over the five-week period is `r accel_df %>% summarize(mean_activity = mean(activity_count, na.rm = TRUE))`. 

Create a total activity variable for each day and create a table. 

```{r}
accel_df %>% 
  group_by(day_id, week, day, weekday_weekend) %>% 
  summarize(total_activity = sum(activity_count)) %>% 
  knitr::kable()
```

It is difficult to identify overall trends with this table but we can see that the minimum total activity, a significant decrease from the average total activity count, occurred on Saturdays. 

Make a single-panel plot. 

```{r}
accel_df %>% 
  ggplot(aes(x = activity_minute, y = activity_count, group = day_id, color = day)) +
  geom_line(alpha = .4) +
  geom_smooth(aes(group = day)) +
  labs(
    title = "24-Hour Activity Time", 
    x = "Minute of 24-Hour Day", 
    y = "Activity Count",
    caption = "Accelerometer data from participant at Columbia University Medical Center") + 
  theme_set(theme_minimal())
```

From this graph, we can see that the participant consistently was most active early in the morning, with the activity level elevated on Friday. There is a also a peak in activity on Sunday in the evening. The participant was least active around 10am on all days.  

## Problem 3 

```{r}
library(p8105.datasets)
data("ny_noaa")
```

This dataset contains `r nrow(ny_noaa)` rows and `r ncol(ny_noaa)` columns. The variables of the dataset include: `r variable.names(ny_noaa)`. There are many missing values for the precipitation, snowfall, snowdepth, and temperature variables. 

Clean ny_noaa dataset.

```{r}
ny_noaa %>% 
  separate(date, c("year", "month", "day"), sep = "-", convert = TRUE) %>% mutate(tmin = as.numeric(tmin)) %>% 
mutate(tmax = as.numeric(tmax)) %>% 
  mutate(tmin = tmin/10) %>% 
  mutate(tmax = tmax/10) 
```

Make a two-panel plot showing the average max temp in Jan and July.


