---
title: "Homework 3"
author: Allison Stewart
output: github_document
---

This is my solution for Homework 3. 

```{r}
library(tidyverse)
library(readxl)
library(dplyr)
library(ggplot2)
library(patchwork)
library(ggridges)
library(hexbin)
```

## Problem 1 

```{r}
library(p8105.datasets)
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns. Observations are at the level of items in orders by user. There are also item variables -- name, aisle, department, and some numeric codes. 

How many aisles, and which are most items from? 

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Make a plot! 

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(
    aisle = factor(aisle), 
    aisle = fct_reorder(aisle,n)
  ) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1))
```

Make a table! 

```{r}
instacart %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetable fruits")) %>% 
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(aisle, rank) %>% 
  knitr::kable()
```

Apples vs. ice cream 

```{r}
instacart %>% 
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>% 
  group_by(product_name, order_dow) %>% 
  summarize(mean_hour = mean(order_hour_of_day)) %>% 
  pivot_wider(
    names_from = order_dow, 
    values_from = mean_hour
  )
```

## Problem 2 

```{r}
accel_df = 
  read.csv("./accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440, 
    names_to = "activity_minute", 
    values_to = "activity_count") %>% 
  relocate(day_id) %>% 
  mutate("weekday_weekend" = day) %>% 
  mutate(weekday_weekend = recode(weekday_weekend, Monday = "Weekday", Tuesday = "Weekday", Wednesday = "Weekday", Thursday = "Weekday", Friday = "Weekday", Saturday = "Weekend", Sunday = "Weekend")) %>% 
  mutate_at(vars(activity_minute), as.factor) %>% 
  mutate_at(vars(day), as.factor) %>% 
  mutate_at(vars(weekday_weekend), as.factor)
```

This dataset contains five weeks of accelerometer data collected on a 63-year-old male who was admitted to the Columbia University Medical Center and diagnosed with congestive heart failure. The variables in the dataset are `r variable.names(accel_df)`. The activity_minute variable represents each minute of a 24-hour day starting at midnight, with the respective activity counts for each minute shown by the activity_count variable. The final table contains `r nrow(accel_df)` rows and `r ncol(accel_df)` columns. The mean activity count over the five-week period is `r accel_df %>% summarize(mean_activity = mean(activity_count, na.rm = TRUE))`. 

Create a total activity variable for each day and create a table. 

```{r}
accel_df %>% 
  group_by(day_id, week, day, weekday_weekend) %>% 
  summarize(total_activity = sum(activity_count)) %>% 
  knitr::kable()
```

It is difficult to identify overall trends with this table but we can see that the minimum total activity, a significant decrease from the average total activity count, occurred on Saturdays. 

Make a single-panel plot. 

```{r}
accel_df %>% 
  ggplot(aes(x = activity_minute, y = activity_count, group = day_id, color = day)) +
  geom_line(alpha = .4) +
  geom_smooth(aes(group = day)) +
  labs(
    title = "24-Hour Activity Time", 
    x = "Minute of 24-Hour Day", 
    y = "Activity Count",
    caption = "Accelerometer data from participant at Columbia University Medical Center") + 
  theme_set(theme_minimal())
```

From this graph, we can see that the participant consistently was most active early in the morning, with the activity level elevated on Friday. There is a also a peak in activity on Sunday in the evening. The participant was least active around 10am on all days.  

## Problem 3 

```{r}
library(p8105.datasets)
data("ny_noaa")
```

This dataset contains `r nrow(ny_noaa)` rows and `r ncol(ny_noaa)` columns. The variables of the dataset include: `r variable.names(ny_noaa)`. There are many missing values for the precipitation, snowfall, snowdepth, and temperature variables. 

Clean and mutate ny_noaa dataset.

```{r}
weather_df = 
  ny_noaa %>% 
  separate(date, c("year", "month", "day"), sep = "-", convert = TRUE) %>% mutate(tmin = as.numeric(tmin)) %>% 
mutate(tmax = as.numeric(tmax)) %>% 
  mutate(tmin = tmin/10) %>% 
  mutate(tmax = tmax/10)
```

Make a two-panel plot showing the average max temp in Jan and July.

```{r}
tmax_weather = 
  weather_df %>% 
  filter(month %in% c("1", "7")) %>% 
  group_by(id, year, month) %>% 
  summarize(mean_tmax = mean(tmax, na.rm = TRUE)) 

ggplot(tmax_weather, aes(x = year, y = mean_tmax)) + geom_point() + facet_grid(~ month) +
  labs(
    title = "Mean average temperature for January and July across stations and years", 
    x = "year", 
    y = "average maximum temperature (C)")
```

From these plots, we can see that there is more variability in average maximum temperature during the month of January as compared to July. The average maximum temperature for July is about 25 degrees Celsius higher than that of January. We can see a few outliers, including a maximum temperature of 14 degrees Celsius in July of 1998 and a maximum temperature of about 4 degrees Celsius in January of 1982. 

Plot tmax vs tmin and snowfall distribution. 

```{r}
p3_plot_2 = 
weather_df %>% 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_hex() + 
 labs(
    title = "Maximum Temperature vs. Minimum Temperature", 
    x = "Minimum Temperature (C)", 
    y = "Maximum Temperature (C)")

p3_plot_3 = 
  weather_df %>% 
  filter(snow > 0 & snow < 100) %>% 
  ggplot(aes(x = year, y = snow, group = year)) +
  geom_violin() +
  labs(
    title = "Snowfall Distribution", 
    x = "Year", 
    y = "Snowfall Distribution(mm)")

(p3_plot_2 + p3_plot_3)
```

The Maximum Temperature vs. Minimum Temperature plot shows a pattern in which maximum temperature temperature as minimum temperature increases. There are several outliers, notably a couple data points that represent a minimum and maximum temperature of 60 degrees Celsius. We the most observations to be concentrated between 0 and 30 degrees celsius. 

The Snowfall Distribution plot indicates a slight increase in median rainfall from 1980 to 1995 and then a decrease from 1995 to 2010. 
